---
title: "Is Correlation Enough?"
description: |
  Questioning whether the mathematical construct: "correlation" is the
  best measure of determining substitute securities
author: Brandon Farr
date: 02-07-2021
categories: ['tidyquant', 'tibbletime', 'finance']
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, code_folding = NULL)
```

**Short-comings of previous analysis**

- we were lucky that the example dataset showed extreme "non-stationarity"
  in correlation between stocks
- this naturally raises the question of whether correlation (not to mention
  which correlation: full data-set, rolling window, ...) is the appropriate
  measure to use in seeking substitutes

```{r}
#| echo: true

library(tidyverse)  # for tidy/dplyr work
library(rvest)      # for web-scraping
library(tidyquant)  # for quant work (tiingo pricing, tq_mutate ...)
library(hrbrthemes) # preferred theming base
library(oqthemes)   # oq colors

```

Once this chunk has been run successfully, just load the data from the local
csv`s: `sp500.csv` and `peer_pricing.csv`.


Again, load the pre-stored data, and calculate returns.

```{r}
#| label: load-local-data

peer_pricing_tbl <- read_csv("data/peer_pricing.csv")

glimpse(peer_pricing_tbl)

# calculate returns as well
peer_returns_tbl <- peer_pricing_tbl %>%
  group_by(symbol) %>%
  tq_transmute(adjusted, periodReturn, period = "daily", col_rename = "return")

```


In a minor adjustment from last-time, begin the creation of the peers table with a direct call to `left_join`, making the two data sets more explicit versus the pipe version.

```{r}

aapl_peer_returns_tbl <-
  left_join(
    # desc
    peer_returns_tbl %>%
      filter(symbol == "AAPL") %>%
      select(symbol, date, return),
    
    # desc
    peer_returns_tbl %>%
      filter(symbol != "AAPL") %>%
      select(symbol, date, return),
    by = "date",
    suffix = c("_1", "_2")
  ) %>%
  select(date, matches("symbol"), everything()) %>%
  arrange(symbol_2, date) %>%
  filter(!(return_1 == 0 & return_2 == 0))  # remove days where returns == 0

```

As we saw last time, even the "smoothed" representation of the 30-day
rolling correlation exhibits significant non-stationarity.

```{r}
#| code-fold: show

library(tibbletime)

cor_roll <- rollify(~cor(.x, .y), window = 30)

aapl_cor_tbl <-
  aapl_peer_returns_tbl %>%
  group_by(symbol_2) %>%
  mutate(cor = cor_roll(return_1, return_2))

aapl_cor_tbl %>%
  ggplot(aes(x = date, y = cor, color = symbol_2)) +
  geom_hline(yintercept = 0, color = "gray40") +
  geom_smooth(se = FALSE) + # default loess smoothing
  ylim(-1, 1) +
  labs(
    title = "AAPL Stock Correlation",
    subtitle = "with GICS sub-industry peers",
    caption = "30-trading day rolling windows",
    x = NULL, y = NULL, color = NULL
  ) +
  theme_ipsum() +
  theme(legend.position = "bottom") +
  scale_color_oq() +
  # scale_x_date(expand = expansion(mult = c(0, 0))) +
  # for legend to have 1 row
  guides(col = guide_legend(nrow = 1))

```

## What is a substitute

When looking for a *substitute* for portfolio purposes, we are seeking:

- a stock that will perform in line with a base security over a given time-frame

We know that period returns are the compounded result of daily returns. So,
maybe the correlation over the data time frame of the period returns (here
30 days) is a better measure of a good substitute.

Or, perhaps the "active" return of the two securities (base and substitute)
through time is a better, simpler measure as it measure the actual gain/loss
that would be experienced by substituting.

```{r}

period_return = function(return) { cumprod(1 + return) %>% last() - 1}

pr_roll <- rollify(~period_return(.x), window = 30)

aapl_pr_tbl <- aapl_peer_returns_tbl %>%
  group_by(symbol_2) %>%
  mutate(
    pr_1 = pr_roll(return_1),
    pr_2 = pr_roll(return_2)
  )


aapl_pr_tbl %>%
  group_by(symbol_2) %>%
  mutate(cor = cor_roll(pr_1, pr_2)) %>%
  ggplot(aes(x = date, y = cor, color = symbol_2)) +
  geom_hline(yintercept = 0, color = "gray40") +
  geom_smooth(se = FALSE) + # default loess smoothing
  ylim(-1, 1) +
  labs(
    title = "AAPL Stock Correlation",
    subtitle = "with GICS sub-industry peers",
    caption = "30-trading day rolling windows",
    x = NULL, y = NULL, color = NULL
  ) +
  theme_ipsum() +
  theme(legend.position = "bottom") +
  scale_color_oq() +
  # scale_x_date(expand = expansion(mult = c(0, 0))) +
  # for legend to have 1 row
  guides(col = guide_legend(nrow = 1))
```

